### 2019-10-11 introduction.tex
intro

\squishlist
\item {\em Overview modality.}
\saj{
One approach for an overview is to simply add a
zoomed out version of the entire sheet as a pane on the side,
as in popular presentation software like 
Microsoft PowerPoint or text editors
like Sublime Text.
\newsaj{The zoomed out overview
displays the entire data 
at a lower magnification scale.} 
Unfortunately, this approach
would not suffice for a spreadsheet.
\new{An overview should provide a
comprehensible big picture view:
for a \newsaj{spreadsheet of numbers, text, or formulae}, 
when zooming out beyond a point, 
\newsaj{an overview displayed at 
such low magnification scale} 
would be unreadable.}
Another approach, adopted by map tools like the
early versions of Google maps,
is to use the overview
to provide a global context of the user's current location
being displayed in the zoomed in detailed view~\cite{cockburn2009review}.
While the overview remains static, users can perform semantic zooming operations~\cite{perlin1993pad} on the detailed
view which allows objects to be represented differently at different scales. Since spreadsheets already display the raw data,
zooming into and out of a detailed view consisting of
this raw data is not meaningful. 
\new{How do we design an overview to dynamically 
change as users 
seek a more fine-grained or coarse-grained view 
of the overall structure of the data?} 
}

\item {\em Construction of the overview.}
\saj{Given a spreadsheet
with many rows, 
one approach to constructing a dynamic overview 
is mapping rows of data to high level groups, 
similar to online maps. 
In online maps, cities are grouped into states 
and states are grouped into countries, 
forming a multi-granularity hierarchy. 
How do we automatically group spreadsheet 
rows together in a similar ``meaningful''
way
such that this grouping applies to all 
data types, including strings and numbers? 
If the automatically generated grouping is 
not semantically meaningful, how do we allow the users
to customize the grouping modality? 
How do we facilitate interactions that enable 
users to view the overview at multiple granularities?}

\item {\em Operations on the overview.}
\saj{Following the construction of a dynamic overview, 
the next challenge is to design simple interactions 
that achieve similar outcomes as scrolling and steering.
For example, an alternative to scrolling 
can be to leverage the groups of the overview 
to access the rows mapped to that group. 
\newsaj{As the granularity of the dynamic overview 
changes, how do we efficiently update the mapping 
from spreadsheet rows to the 
finer or coarser groups 
so that scrolling remains seamless? 
Similarly, how do we leverage
the overview to steer spreadsheet data (i.e., select a range of data)
for formula computation? How do we present the results of
the formula within the overview? One approach can be to adopt
the pivot table-like presentation of results. Within their summary view, 
pivot tables display 
aggregate formula results (\eg \code{SUM}, \code{COUNT}) 
alongside each group. 
However, unlike pivot tables, users can explore the overview at
multiple-granularities. As the granularity changes,
the grouping of rows also changes; 
making the previous formula results 
inconsistent with the new groups.
How do we 
recompute the results of a formula 
as the granularity changes 
without requiring the users to 
reissue the formula from scratch?}}
\agp{the last three questions are fairly confusing and assume that you already know what NOAH does.. can we simplify?}

\item {\em Seamless integration as a plug-in.}
\newsaj{Finally, how can we 
design an overview with a generic set of features,
that can be integrated
with any existing spreadsheet tool on any 
dataset without impacting
existing functionalities or look-and-feel?
How do we ensure that the interactions 
supported by the overview 
are 
consistent with
traditional spreadsheet semantics,
and complement existing spreadsheet interactions?
How to we enable coordinated interactions across both views, \ie 
the overview and the raw spreadsheet, such that they
remain consistent at any given time?}
\squishend
\agp{A general comment: can we ensure that either we use
questions for all of the items above or none of them for uniformity?}

\stitle{\noah: \new{a navigation plug-in for spreadsheets}.} We address the aforementioned challenges
in \noah,
an in-situ
\textbf{\underline{n}}avigation interface
for \textbf{\underline{o}}verviewing
and \textbf{\underline{a}}nalyzing
spreadsheet data \textbf{\underline{h}}olistically.
\noah is constructed as a plugin to
an existing spreadsheet tool,
{\scshape DataSpread}~\cite{dataspread},
an open-source scalable web-based spreadsheet.
While \noah's design is not tied to {\sc DataSpread},
we opted not to use other popular spreadsheet
tools like Google Sheets and Microsoft Excel because they
are closed source.
Figure~\ref{fig:ux} shows a snapshot of \noah.
When the user chooses to explore
the data by a specific attribute,
\saj{a multi-granularity overview is constructed and displayed
within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a). Users can zoom into or out of the overview to obtain a \new{fine or coarse-}grained 
perspective of the data distribution. 
The distribution at each granularity 
is captured by a histogram,
enabling users to assimilate the data \new{via summary
statistics}. 
Each bin (group) of the histogram is 
mapped to a collection of rows in the spreadsheet.} 
Cumbersome scrolling operations are eliminated in favor
of a few clicks on the overview interface. Instead of steering to analyze the data,
users can issue formulae
on the overview with interactions similar to pivot table construction,
and view results on a separate {\em aggregate column},
alongside the overview (Figure~\ref{fig:ux}b). 
In this manner,
users can issue formulae on different
subsets of the data while remaining on the same screen,
reducing visual discontinuity.
\noah ensures that there is coordination between
the overview and the spreadsheet:
for example, panning and zooming on the overview
are reflected on the spreadsheet by displaying
the spreadsheet data corresponding to the bin
currently in focus in the overview.
Finally, \noah automatically creates
contextual and historical information
(Figure~\ref{fig:ux}d and~\ref{fig:ux}e)
while displaying visual cues
(Figure~\ref{fig:ux}f and~\ref{fig:ux}g)
so that users don't lose context during navigation.

The primary contribution of our work is twofold:
\squishlist
\item We formalize the design of a general
navigation (overview+detail) interface
for exploration and analysis of large spreadsheets.
We realize this design in the form of \noah,
a plugin to an existing web-based spreadsheet tool,
ensuring that interactions supported by \noah
complement existing spreadsheet operations.
\item We conduct a user study to evaluate
the benefits and limitations of this plugin.
\newsaj{The study required users to
perform tasks that were representative of
popular spreadsheet operations,
using both Microsoft Excel and \noah.}
\saj{The study shows that compared to Microsoft Excel, participants were able to complete spreadsheet
tasks involving navigation 
correctly and quickly in \noah. Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} \agp{mention how these tasks were representative or general}
\squishend

>>>>

\squishlist
\item {\em Overview modality.}
\new{
One could simply add a
zoomed out version of the entire sheet as a pane on the side
to create a spreadsheet overview,
as in popular presentation software like 
Microsoft PowerPoint, or text editors
like Sublime Text.
The zoomed out overview
would 
display the data 
at a lower magnification. 
Unfortunately, this approach
would not suffice for a spreadsheet.
An overview should provide a
comprehensible big picture view;
for a spreadsheet of numbers, text, or formulae, 
when zooming out beyond a point, 
an overview displayed at 
such low magnification 
would be unreadable.
Another approach, adopted by map tools like the
early versions of Google maps,
is to use the overview
to provide a global context of the user's current location
currently displayed in the zoomed-in detailed view~\cite{cockburn2009review}.
While the overview remains static, users can perform semantic zooming operations~\cite{perlin1993pad} on the detailed
view which allows objects to be represented differently at different scales. 
Since spreadsheets already display the raw data,
zooming into and out of a detailed view consisting of
this raw data is not meaningful. 
How do we design an overview to dynamically 
change as users 
seek a more fine-grained or coarse-grained view 
of the overall structure of the data?
}

\item {\em Construction of the overview.}
\new{Given a spreadsheet
with many rows, 
one approach to constructing a dynamic overview 
is mapping rows of data to high level groups, 
similar to online maps. 
In online maps, cities are grouped into states 
and states are grouped into countries, 
forming a multi-granularity hierarchy. 
How do we automatically group spreadsheet 
rows together in a similar ``meaningful''
way
such that this grouping applies to all 
data types, including strings and numbers? 
If the automatically generated grouping is 
not semantically meaningful, how do we allow the users
to customize the grouping modality? 
How do we facilitate interactions that enable 
users to view the overview at multiple granularities?}

\item {\em Operations on the overview.}
\new{Following the construction of a dynamic overview, 
the next challenge is to design simple interactions 
that achieve similar outcomes as scrolling and steering.
For example, an alternative to scrolling 
can be to leverage the groups of the overview 
to access the rows mapped to that group. 
As the granularity of the dynamic overview 
changes, how do we efficiently update the mapping 
from spreadsheet rows to the 
finer or coarser groups 
so that scrolling remains seamless? 
Similarly, how do we leverage
the overview to steer spreadsheet data (i.e., select a range of data)
for formula computation? How do we present the results of
the formula within the overview? One approach can be to adopt
the pivot table-like presentation of results. 
Within their summary view, 
pivot tables display 
aggregate formula results (\eg \code{SUM}, \code{COUNT}) 
alongside each group. 
However, unlike pivot tables, users can explore the overview at
multiple granularities. As the granularity changes,
the grouping of rows also changes; 
making the previous formula results 
inconsistent with the new groups.
How do we 
recompute the results of a formula 
in a convenient manner
as the granularity changes 
without requiring the users to 
reissue the formula from scratch?}

\item {\em Seamless integration as a plug-in.}
\new{Finally, how can we 
design an overview with a generic set of features,
that can be integrated
with any existing spreadsheet tool,
operating on any 
dataset, without impacting
existing functionalities or look-and-feel?
How do we ensure that the interactions 
supported by the overview 
are 
consistent with
traditional spreadsheet semantics,
and complement existing spreadsheet interactions?
How to we enable coordinated interactions across both views, \ie 
the overview and the raw spreadsheet, such that they
remain consistent at any given time?}
\squishend

\stitle{\noah: \new{a navigation plug-in for spreadsheets}.} We address the aforementioned challenges
in \noah,
an in-situ
\textbf{\underline{n}}avigation interface
for \textbf{\underline{o}}verviewing
and \textbf{\underline{a}}nalyzing
spreadsheet data \textbf{\underline{h}}olistically.
\noah is constructed as a plugin to
an existing spreadsheet tool,
{\scshape DataSpread}~\cite{dataspread},
an open-source scalable web-based spreadsheet.
While \noah's design is not tied to {\sc DataSpread},
we opted not to use other popular spreadsheet
tools like Google Sheets and Microsoft Excel because they
are closed source.
Figure~\ref{fig:ux} shows a snapshot of \noah.
When the user chooses to explore
the data by a specific attribute,
\new{a multi-granularity overview is constructed and displayed
within \noah, next to the raw spreadsheet data (Figure~\ref{fig:ux}a). Users can zoom into or out of the overview to obtain a fine or coarse-grained 
perspective of the data distribution. 
The distribution at each granularity 
is captured by a histogram,
enabling users to assimilate the data \new{via summary
statistics}. 
Each bin (group) of the histogram is 
mapped to a collection of rows in the spreadsheet.} 
Cumbersome scrolling operations are eliminated in favor
of a few clicks on the overview interface. Instead of steering to analyze the data,
users can issue formulae
on the overview with interactions similar to pivot table construction,
and view results on a separate {\em aggregate column},
alongside the overview (Figure~\ref{fig:ux}b). 
In this manner,
users can issue formulae on different
subsets of the data while remaining on the same screen,
reducing visual discontinuity.
\noah ensures that there is coordination between
the overview and the spreadsheet:
for example, panning and zooming on the overview
are reflected on the spreadsheet by displaying
the spreadsheet data corresponding to the bin
currently in focus in the overview.
Finally, \noah automatically creates
contextual and historical information
(Figure~\ref{fig:ux}d and~\ref{fig:ux}e)
while displaying visual cues
(Figure~\ref{fig:ux}f and~\ref{fig:ux}g)
so that users don't lose context during navigation.

The primary contribution of our work is twofold:
\squishlist
\item We formalize the design of a general
navigation (overview+detail) interface
for exploration and analysis of large spreadsheets.
We realize this design in the form of \noah,
a plugin to a spreadsheet tool,
ensuring that interactions supported by \noah
complement existing spreadsheet operations.
\item We conduct a user study to evaluate
the benefits and limitations of this plugin.
\new{The study required users to
perform tasks that were representative of
popular spreadsheet operations,
using both Microsoft Excel and \noah.
The study shows that compared to Microsoft Excel, 
participants were able to complete spreadsheet
navigation tasks 
correctly and quickly in \noah. Participants made {\bf $2.5\times$ fewer} mistakes while being {\bf $2\times$ faster} with a \noah-integrated spreadsheet than with Excel.} 
\squishend
